<font face="微软雅黑">

---

# **第七章：生态数据的空间分析**

---

本章内容简介：

- 7.1 目标
- 7.2 空间结构与空间分析简介
- 7.3 多变量趋势面分析(Multivariate Trend-Surface Analysis)
- 7.4 基于特征向量的空间变量与空间模型(Eigenvector-Based Spatial Variables and Spatial Modelling)
- 7.5 另一种空间结构的观察方法: 多尺度排序(Multiscale Ordination, MSO)
- 7.6 多元变量方差分析中的无重复时空交互检验(Space-Time Interaction Test in Multivariate ANOVA, Without Replicates)
- 7.7 结论

---

## 7.1 目标


---

## 7.2 空间结构与空间分析简介

### 7.2.1 简介

### 7.2.2 诱导空间依赖与空间自相关(Induced Spatial Dependence and Spatial Autocorrelation)

### 7.2.3 空间尺度(Spatial Scale)

### 7.2.4 空间异质性(Spatial Heterogeneity)

### 7.2.5 空间相关或相关函数与空间相关图(Spatial Correlation or Autocorrelation Functions and Spatial Correlograms)


```{r}
# Load packages, functions and data ===============================
library(ape)
library(spdep)
library(ade4)
library(adegraphics)
library(adespatial)
library(vegan)


# Source additional functions that will be used later in this
# Chapter. Our scripts assume that files to be read are in
# the working directory.
source("Functions/plot.links.R")
source("Functions/sr.value.R")
source("Functions/quickMEM.R")
source("Functions/scalog.R")

# Load the oribatid mite data. The file mite.Rdata is assumed 
# to be in the working directory.
load("Data/mite.RData")

# Transform the data
mite.h <- decostand (mite, "hellinger")
mite.xy.c <- scale(mite.xy, center = TRUE, scale = FALSE)



# Univariate spatial correlogram (based on Moran's I) =============

# Search for neighbours of all points within a radius of 0.7 m
# and multiples (i.e., 0 to 0.7 m, 0.7 to 1.4 m and so on).
# The points do not form a connected graph at 0.7 m.
 
plot.links(mite.xy, thresh = 0.7)
nb1 <- dnearneigh(as.matrix(mite.xy), 0, 0.7)
summary(nb1)

# Correlogram of substrate density
subs.dens <- mite.env[ ,1]
subs.correlog <- 
  sp.correlogram(nb1, 
                 subs.dens, 
                 order = 14, 
                 method = "I", 
	     zero.policy = TRUE)
print(subs.correlog, p.adj.method = "holm")
 
plot(subs.correlog)



# Mantel correlogram of the oribatid mite data ====================

# The species data are first detrended; see Sect. 7.3
mite.h.det <- resid(lm(as.matrix(mite.h) ~ ., data = mite.xy))
mite.h.D1 <- dist(mite.h.det)
(mite.correlog <- 
  mantel.correlog(mite.h.D1, 
                  XY = mite.xy, 
                  nperm = 999))
summary(mite.correlog)

# Number of classes
mite.correlog$n.class # or: mite.correlog[2]
# Break points
mite.correlog$break.pts # or: mite.correlog[3]

# Plot the Mantel correlogram
 
plot(mite.correlog)

```


### 7.2.6 空间相关性存在的检验: 条件(Testing for the Presence of Spatial Correlation: Conditions)

### 7.2.7 空间结构建模(Modelling Spatial Structures)


---

## 7.3 多变量趋势面分析(Multivariate Trend-Surface Analysis)

### 7.3.1 简介

### 7.3.2 趋势面分析的实际应用(In practice)



```{r}

# Trend-surface analysis ==========================================

## Simple models on a square, regularly sampled surface

# Construct and plot a 10 by 10 grid
xygrid <- expand.grid(1:10, 1:10)
 
plot(xygrid)

# Centring
xygrid.c <- scale(xygrid, scale = FALSE)

# Create and plot some first, second and third-degree functions 
# of X and Y
X <- xygrid.c[ ,1]
Y <- xygrid.c[ ,2]
XY <- X + Y
XY2 <- X ^ 2  +  Y ^ 2
XY3 <- X ^ 2 - X * Y - Y ^ 2
XY4 <- X + Y  +  X ^ 2  +  X * Y  +  Y ^ 2
XY5 <- X ^ 3  +  Y ^ 3
XY6 <- X ^ 3  +  X ^ 2 * Y  +  X * Y ^ 2  +  Y ^ 3
XY7 <- X  +  Y  +  X ^ 2  +  X * Y  +  Y ^ 2  +  X ^ 3  
          +  X ^ 2 * Y  +  X * Y ^ 2  +  Y ^ 3
xy3deg <- cbind(X, Y, XY, XY2, XY3, XY4, XY5, XY6, XY7)

 
s.value(xygrid, xy3deg, symbol = "circle")
     # Try other combinations, for instance with minus signs or
     # with coefficients not equal to 1.


## Trend-surface analysis of the mite data

# Computation of a raw (non-orthogonal) third-degree polynomial 
# function on the previously centred X-Y coordinates
mite.poly <- poly(as.matrix(mite.xy.c), degree = 3, raw = TRUE)
colnames(mite.poly) <- 
    c("X", "X2", "X3", "Y", "XY", "X2Y", "Y2", "XY2", "Y3")

# RDA with all 9 polynomial terms
(mite.trend.rda <- rda(mite.h ~ ., 
                       data = as.data.frame(mite.poly)))

# Computation of the adjusted R^2
(R2adj.poly <- RsquareAdj(mite.trend.rda)$adj.r.squared)

# RDA using a third-degree orthogonal polynomial of the geographic 
# coordinates
mite.poly.ortho <- poly(as.matrix(mite.xy), degree = 3)
colnames(mite.poly.ortho) <- 
    c("X", "X2", "X3", "Y", "XY", "X2Y", "Y2", "XY2", "Y3")
(mite.trend.rda.ortho <- 
  rda(mite.h ~ ., 
      data = as.data.frame(mite.poly.ortho)))
(R2adj.poly2 <- RsquareAdj(mite.trend.rda.ortho)$adj.r.squared)

# Forward selection using Blanchet et al. (2008a) double stopping 
# criterion
(mite.trend.fwd <- 
  forward.sel(mite.h, mite.poly.ortho, adjR2thresh = R2adj.poly2))

# New RDA using the 6 terms retained
(mite.trend.rda2 <- rda(mite.h ~ ., 
          data = as.data.frame(mite.poly)[ ,mite.trend.fwd[ ,2]]))

# Overall test and test of the canonical axes
anova(mite.trend.rda2)
anova(mite.trend.rda2, by = "axis")

# Plot of the three independent significant spatial structures
# (canonical axes) plus the fourth (p-value around 0.06). 
mite.trend.fit <- 
  scores(mite.trend.rda2, 
         choices = 1:4, 
         display = "lc", 
         scaling = 1)
 
s.value(mite.xy, mite.trend.fit, symbol = "circle")

# Other plotting code with homemade function sr.value:
# dev.new(title = "Mite Trend Surface Analysis")
# par(mfrow = c(1,3))
# sr.value(mite.xy,mite.trend.fit[ ,1])
# sr.value(mite.xy,mite.trend.fit[ ,2])
# sr.value(mite.xy,mite.trend.fit[ ,3])

# -----------------------------------------------------------------
  # If you want to construct a raw polynomial function directly
  # within the rda call, here is the syntax (2nd degree):
  # mite.trend.rda <- 
  #            rda(mite.h ~ Xm + Ym + I(Xm^2) + I(Xm*Ym) + I(Ym^2))
  # Notice how squared variables and product variables are 
  # requested to be treated "as they are" by function I(). 
  # Otherwise R would consider them as ANOVA terms.
# -----------------------------------------------------------------

```


---

## 7.4 基于特征向量的空间变量与空间模型(Eigenvector-Based Spatial Variables and Spatial Modelling)

### 7.4.1 简介

### 7.4.2 基于距离的 Moran 特征向量图(Distance-Based Moran's Eigenvector Maps, dbMEM) 和 相邻矩阵主坐标分析(Principal Coordinates of Neighbour Matrices, PCNM)


#### 7.4.2.2 常规采样设计的 dbMEM 变量 (Regular Sampling Designs)

```{r}
## dbMEM analysis (artificial data)

# 1. One-dimensional sampling: transect with 100 equispaced points.
#    The distance between adjacent points is 1. Function dbmem()
#    automatically computes the threshold value, 1 in this case.

# Generate transect points
tr100 <- 1 : 100

# Creation of the dbMEM eigenfunctions with Moran's I corresponding
# to positive spatial correlation: argument
# MEM.autocor = "positive" which is the default
tr100.dbmem.tmp <- dbmem(tr100, silent = FALSE)
tr100.dbmem <- as.data.frame(tr100.dbmem.tmp)
# Display the eigenvalues
attributes(tr100.dbmem.tmp)$values
# Number of (positive) eigenvalues
length(attributes(tr100.dbmem.tmp)$values)

# Plot some dbMEM variables modelling positive spatial correlation
#      along a transect (Fig. 7.3).
 
# par(mfrow = c(4, 2))
somedbmem <- c(1, 2, 4, 8, 15, 20, 30, 40)
for(i in 1:length(somedbmem)){
	plot(tr100.dbmem[ ,somedbmem[i]], 
	     type = "l", 
	     xlab = "X coordinate", 
	     ylab = c("dbMEM", somedbmem[i]))
}

# 2. Two-dimensional sampling: grid of equispaced points with 
#    smallest distance between points equal to 1 (Fig. 7.4).

# Generate grid point coordinates
xygrid2 <- expand.grid(1:20, 1:20)

# Creation of the dbMEM eigenfunctions with positive Moran's I 
xygrid2.dbmem.tmp <- dbmem(xygrid2)
xygrid2.dbmem <- as.data.frame(xygrid2.dbmem.tmp)
# Count the eigenvalues
length(attributes(xygrid2.dbmem.tmp)$values)

# Plot some dbMEM variables using s.value {adegraphics}
 
somedbmem2 <- c(1, 2, 5, 10, 20, 40, 80, 120, 189)
s.value(xygrid2, xygrid2.dbmem[ ,somedbmem2], 
  method = "color", 
  symbol = "circle", 
  ppoints.cex = 0.5
)

# Plot some dbMEM variables (slow)
# Using homemade function sr.value
# dev.new(title = "dbMEM variables (grid)", noRStudioGD = TRUE)
# par(mfrow = c(3,3))
# somedbmem2 <- c(1, 2, 5, 10, 20, 40, 80, 120, 189)
# for(i in 1:length(somedbmem2))
# {
#    sr.value(xygrid2, xygrid2.dbmem[ ,somedbmem2[i]], 
#             method = "greylevel", 
# 	  csize = 0.65,
#             sub = somedbmem2[i], 
#             csub = 1.8)
# }

```


#### 7.4.2.3 甲螨数据集的基于距离的 Moran 特征向量图分析(Distance-Based Moran's Eigenvector Maps, dbMEM)

```{r}
## dbMEM analysis of the oribatid mite data

# Is there a linear trend in the mite data?
anova(rda(mite.h, mite.xy))	# Result: significant trend
# Computation of linearly detrended mite data
mite.h.det <- resid(lm(as.matrix(mite.h) ~ ., data = mite.xy))

## Step 1. Construct the matrix of dbMEM variables
mite.dbmem.tmp <- dbmem(mite.xy, silent = FALSE)
mite.dbmem <- as.data.frame(mite.dbmem.tmp)
# Truncation distance used above:
(thr <- give.thresh(dist(mite.xy)))

# Display and count the eigenvalues
attributes(mite.dbmem.tmp)$values
length(attributes(mite.dbmem.tmp)$values)
  # Argument silent = FALSE allows the function to display 
  # the truncation level.

## Step 2. Run the global dbMEM analysis on the *detrended*
##    Hellinger-transformed mite data
(mite.dbmem.rda <- rda(mite.h.det ~ ., mite.dbmem))
anova(mite.dbmem.rda)

## Step 3. Since the R-square is significant, compute the adjusted
##    R2 and run a forward selection of the dbmem variables
(mite.R2a <- RsquareAdj(mite.dbmem.rda)$adj.r.squared)
(mite.dbmem.fwd <- forward.sel(mite.h.det, as.matrix(mite.dbmem), 
	adjR2thresh = mite.R2a))
(nb.sig.dbmem <- nrow(mite.dbmem.fwd))    # Number of signif. dbMEM
# Identity of the significant dbMEM in increasing order
(dbmem.sign <- sort(mite.dbmem.fwd[ ,2]))
# Write the significant dbMEM to a new object
dbmem.red <- mite.dbmem[ ,c(dbmem.sign)]

## Step 4. New dbMEM analysis with 8 significant dbMEM variables
##    Adjusted R-square after forward selection: R2adj = 0.2418
(mite.dbmem.rda2 <- rda(mite.h.det ~ ., data = dbmem.red))
(mite.fwd.R2a <- RsquareAdj(mite.dbmem.rda2)$adj.r.squared)
anova(mite.dbmem.rda2)
(axes.test <- anova(mite.dbmem.rda2, by = "axis"))
# Number of significant axes
(nb.ax <- length(which(axes.test[ ,ncol(axes.test)] <=  0.05)))

## Step 5. Plot the significant canonical axes
mite.rda2.axes <- 
  scores(mite.dbmem.rda2, 
         choices = c(1:nb.ax), 
         display = "lc", 
         scaling = 1)
 
# par(mfrow = c(1,nb.ax))
for(i in 1:nb.ax){
    sr.value(mite.xy, mite.rda2.axes[ ,i], 
             sub = paste("RDA",i), 
             csub = 2)
                 }
				 
# Interpreting the spatial variation: regression of the significant
# canonical axes on the environmental variables, with Shapiro-Wilk 
# normality tests of residuals
mite.rda2.axis1.env <- lm(mite.rda2.axes[ ,1] ~ ., data = mite.env)
shapiro.test(resid(mite.rda2.axis1.env))
summary(mite.rda2.axis1.env)

mite.rda2.axis2.env <- lm(mite.rda2.axes[ ,2] ~ ., data = mite.env)
shapiro.test(resid(mite.rda2.axis2.env))
summary(mite.rda2.axis2.env)

# Depending on the permutation-based p-value the third axis may 
# not be significant. In that case the 3 next  lines of code will
# return error messages.
mite.rda2.axis3.env <- lm(mite.rda2.axes[ ,2] ~ ., data = mite.env)
shapiro.test(resid(mite.rda2.axis3.env))
summary(mite.rda2.axis3.env)

# Scalogram of the variance explained by all dbMEM eigenfunctions, 
# computed with our homemade function scalog()
 
scalog(mite.dbmem.rda)

# Maps of the 8 significant dbMEM variables with the homemade
# function sr.value()
 
# par(mfrow = c(2, 4))
for(i in 1 : ncol(dbmem.red))
{
    sr.value(mite.xy, 
             dbmem.red[ ,i], 
#             sub = paste("dbMEM",i), 
             sub = paste("dbMEM",dbmem.sign[i]), 
             csub = 2)
}

# Maps with s.value()
 
s.value(mite.xy, dbmem.red)



## dbMEM analysis of the mite data - broad scale

(mite.dbmem.broad <- 
  rda(mite.h.det ~ ., data = mite.dbmem[ ,c(1,3,4)]))
anova(mite.dbmem.broad)
(axes.broad <- anova(mite.dbmem.broad, by = "axis"))
# Number of significant axes
(nb.ax.broad <- 
  length(which(axes.broad[ , ncol(axes.broad)] <=  0.05)))

# Plot of the two significant canonical axes
mite.dbmembroad.axes <- 
  scores(mite.dbmem.broad, 
         choices = c(1,2), 
         display = "lc", 
         scaling = 1)
 
# par(mfrow = c(1, 2))
sr.value(mite.xy, mite.dbmembroad.axes[ ,1])
sr.value(mite.xy, mite.dbmembroad.axes[ ,2])

# Interpreting the broad-scaled spatial variation: regression of
# the two significant spatial canonical axes on the environmental
# variables
mite.dbmembroad.ax1.env <-   lm(mite.dbmembroad.axes[ ,1] ~ ., data = mite.env)
summary(mite.dbmembroad.ax1.env)
mite.dbmembroad.ax2.env <-  lm(mite.dbmembroad.axes[ ,2] ~ ., data = mite.env)
summary(mite.dbmembroad.ax2.env)


## dbMEM analysis of the mite data - medium scale

(mite.dbmem.med <-  rda(mite.h.det ~ ., data = mite.dbmem[ ,c(6,7,10,11)]))
anova(mite.dbmem.med)
(axes.med <- anova(mite.dbmem.med, by = "axis"))
# Number of significant axes
(nb.ax.med <- length(which(axes.med[ ,ncol(axes.med)] <=  0.05)))

# Plot of the significant canonical axes
mite.dbmemmed.axes <-  scores(mite.dbmem.med, choices = c(1,2), 
         display = "lc", 
         scaling = 1)
 
# par(mfrow = c(1, 2))
sr.value(mite.xy, mite.dbmemmed.axes[ ,1])
sr.value(mite.xy, mite.dbmemmed.axes[ ,2])

# Interpreting the medium-scaled spatial variation: regression of
# the two significant spatial canonical axes on the environmental
# variables
mite.dbmemmed.ax1.env <-  lm(mite.dbmemmed.axes[ ,1] ~ ., data = mite.env)
summary(mite.dbmemmed.ax1.env)
mite.dbmemmed.ax2.env <-   lm(mite.dbmemmed.axes[ ,2] ~ ., data = mite.env)
summary(mite.dbmemmed.ax2.env)


## dbMEM analysis of the mite data - fine scale

(mite.dbmem.fine <-  rda(mite.h.det ~ ., data = as.data.frame(mite.dbmem[ ,20])))
anova(mite.dbmem.fine)
# Analysis stops here, since the RDA is not significant.


```


#### 7.4.2.4 Hassle-Free dbMEM Analysis: Function quickMEM()

```{r}
mite.dbmem.quick <- quickMEM(mite.h, mite.xy)
summary(mite.dbmem.quick)
# Eigenvalues
mite.dbmem.quick[[2]]  # OR mite.dbmem.quick$eigenvalues
# Results of forward selection
mite.dbmem.quick[[3]]  # OR mite.dbmem.quick$fwd.sel

# Extract and plot RDA results from a quickMEM output (scaling 2)
 
plot(mite.dbmem.quick$RDA, scaling = 2)
sp.scores2 <- 
  scores(mite.dbmem.quick$RDA, 
         choices = 1:2, 
         scaling = 2, 
         display = "sp")
arrows(0, 0, 
  sp.scores2[ ,1] * 0.9, 
  sp.scores2[ ,2] * 0.9, 
  length = 0, 
  lty = 1, 
  col = "red"
)
```



#### 7.4.2.5 结合基于距离的 Moran 特征向量图分析 & 方差分解(Variation Partitioning)


```{r}
# 1. Test trend
mite.XY.rda <- rda(mite.h, mite.xy)
anova(mite.XY.rda)

# 2. Test and forward selection of the environmental variables
# Recode environmental variables 3 to 5 into dummy binary variables
substrate <- model.matrix( ~ mite.env[ ,3])[ ,-1]
shrubs <- model.matrix( ~ mite.env[ ,4])[ ,-1]
topography <- model.matrix( ~ mite.env[ ,5])[ ,-1]
mite.env2 <- cbind(mite.env[ ,1:2], substrate, shrubs, topography)
colnames(mite.env2) <- 
  c("SubsDens", "WatrCont", "Interface", "Litter", "Sphagn1",
    "Sphagn2", "Sphagn3", "Sphagn4", "Shrubs_Many", "Shrubs_None", 
    "topography")
# Forward selection of the environmental variables
mite.env.rda <- rda(mite.h ~., mite.env2)
(mite.env.R2a <- RsquareAdj(mite.env.rda)$adj.r.squared)
mite.env.fwd <- 
  forward.sel(mite.h, mite.env2,
              adjR2thresh = mite.env.R2a, 
              nperm = 9999)
env.sign <- sort(mite.env.fwd$order)
env.red <- mite.env2[ ,c(env.sign)]
colnames(env.red)

# 3. Test and forward selection of the dbMEM variables
# Run the global dbMEM analysis on the *detrended* mite data
mite.det.dbmem.rda <- rda(mite.h.det ~., mite.dbmem)
anova(mite.det.dbmem.rda)
# Since the analysis is significant, compute the adjusted R2
# and run a forward selection of the dbMEM variables
(mite.det.dbmem.R2a <- 
  RsquareAdj(mite.det.dbmem.rda)$adj.r.squared)
(mite.det.dbmem.fwd <- 
  forward.sel(mite.h.det, 
              as.matrix(mite.dbmem), 
	  adjR2thresh = mite.det.dbmem.R2a))
# Number of significant dbMEM
(nb.sig.dbmem <- nrow(mite.det.dbmem.fwd))
# Identify the significant dbMEM sorted in increasing order
(dbmem.sign <- sort(mite.det.dbmem.fwd$order))
# Write the significant dbMEM to a new object (reduced set)
dbmem.red <- mite.dbmem[ ,c(dbmem.sign)]

# 4. Arbitrarily split the significant dbMEM into broad and 
#    fine scale
# Broad scale: dbMEM 1, 3, 4, 6, 7
dbmem.broad <- dbmem.red[ , 1 : 5]
# Fine scale: dbMEM 10, 11, 20
dbmem.fine <- dbmem.red[ , 6 : 8]

## 5. Mite - environment - trend - dbMEM variation partitioning
(mite.varpart <- 
  varpart(mite.h, env.red, mite.xy, dbmem.broad, dbmem.fine))
 
# Show the symbols of the fractions and plot their values
# par(mfrow = c(1,2))
showvarparts(4, bg = c("red", "blue", "yellow", "green"))
plot(mite.varpart, 
     digits = 2, 
     bg = c("red", "blue", "yellow", "green")
) 
# Tests of the unique fractions [a], [b], [c] and [d]
# Fraction [a], pure environmental
anova(
  rda(mite.h, env.red, cbind(mite.xy, dbmem.broad, dbmem.fine))
)
# Fraction [b], pure trend
anova(
  rda(mite.h, mite.xy, cbind(env.red, dbmem.broad, dbmem.fine))
)
# Fraction [c], pure broad scale spatial
anova(rda
  (mite.h, dbmem.broad, cbind(env.red, mite.xy, dbmem.fine))
)
# Fraction [d], pure fine scale spatial
anova(
  rda(mite.h, dbmem.fine, cbind(env.red, mite.xy, dbmem.broad))
)

```



### 7.4.3 更广泛背景下的Moran 特征向量图分析方法(MEM): 地理距离以外的权重(Weights Other than Geographic Distances)


#### 7.4.3.1 简介


#### 7.4.3.2 甲螨数据集的广义 MEM 分析 (的Moran 特征向量图)

```{r}
## MEM analysis of the detrended oribatid mite data

# Selection of an optimal spatial weighting matrix

# 1. Search based on Delaunay triangulation.
#    We use mite.h.det as response data and mite.del as Delaunay
#    triangulation data.
#    No weighting matrix (binary weights only): 1 means connected
#    (easy communication), 0 = not connected (no exchange 
#    possible). Function test.W() selects among the MEM variables 
#    constructed on the basis of the Delaunay triangulation.

# Delaunay triangulation and model selection
(mite.del <- tri2nb(mite.xy)) 
mite.del.res <- 
  test.W(mite.h.det, 
         mite.del, 
         MEM.autocor = "positive")
# Summary of the results for the best model
summary(mite.del.res$best)
# Unadjusted R^2 of the model with the smallest AICc value
(R2.del <- 
  mite.del.res$best$AIC$R2[which.min(mite.del.res$best$AIC$AICc)])
# Adjusted R^2 of the model with the smallest AICc value
RsquareAdj(
  R2.del, 
  n = nrow(mite.h.det), 
  m = which.min(mite.del.res$best$AIC$AICc)
)

# 2. Delaunay triangulation weighted by a function of distance.
#    Distances are ranged to maximum 1, and raised to power y. 
#    After transformation of the distances by function f2, values
#    near 1 are attributed to pairs of sites with easy exchange,
#    values near 0 mean difficult communication.
f2 <- function(D, dmax, y) 
{ 
  1 - (D/dmax)^y 
}

# Largest Euclidean distance on links belonging to the Delaunay 
# triangulation
max.d1 <- max(unlist(nbdists(mite.del, as.matrix(mite.xy)))) 
# Power y is set from 2 to 10
mite.del.f2 <- 
  test.W(mite.h.det, mite.del, 
         MEM.autocor = "positive", 
         f = f2, 
         y = 2:10, 
         dmax = max.d1, 
         xy = as.matrix(mite.xy))
# Unadjusted R^2 of best model
(R2.delW <- 
  mite.del.f2$best$AIC$R2[which.min(mite.del.f2$best$AIC$AICc)])
# Adjusted R^2 of best model
RsquareAdj(
  R2.delW, 
  n = nrow(mite.h.det), 
  m = which.min(mite.del.f2$best$AIC$AICc)
)

# 3a. Connectivity matrix based on a distance (radius around 
#     points)
# Assessment of the relevant distances based on a multivariate 
# variogram of the detrended mite data, with 20 distance classes.
(mite.vario <- variogmultiv(mite.h.det, mite.xy, nclass = 20))
 
plot(
  mite.vario$d, 
  mite.vario$var, 
  ty = 'b', 
  pch = 20, 
  xlab = "Distance", 
  ylab = "C(distance)"
)
# Construction of 10 neighbourhood matrices (class nb)
# Vector of 10 threshold distances
(thresh10 <- seq(give.thresh(dist(mite.xy)), 4, le = 10))
# Create 10 neighbourhood matrices.
# Each matrix contains all connexions with lengths smaller or equal
# to the threshold value
list10nb <- 
  lapply(thresh10, 
         dnearneigh, 
         x = as.matrix(mite.xy), 
         d1 = 0)
# Display an excerpt of the first neighbourhood matrix
print(
  listw2mat(nb2listw(list10nb[[1]], style = "B"))[1:10,1:10], 
  digits = 1
)
# Now we can apply the function test.W() to the 10 neighbourhood
# matrices. There are no weights on the links.
mite.thresh.res <- 
  lapply(list10nb, 
    function(x) test.W(x, 
                       Y = mite.h.det, 
                       MEM.autocor = "positive")
)
# Lowest AICc, best model, threshold distance of best model
mite.thresh.minAIC <- 
  sapply(mite.thresh.res, 
         function(x) min(x$best$AIC$AICc, na.rm = TRUE))
# Smallest AICc (best model among the 10)
min(mite.thresh.minAIC)
# Number of the model among the 10
which.min(mite.thresh.minAIC)
# Truncation threshold (distance)
thresh10[which.min(mite.thresh.minAIC)]

# 3b. Variant: same as above, but connections weighted by the
#     complement of the power of the distances, 1-(d/dmax)^y. 
#     Again, after transformation of the distances by function f2,
#     values near 1 are attributed to pairs of sites with easy
#     exchange, values near 0 mean difficult communication.
mite.thresh.f2 <- 
  lapply(list10nb, 
         function(x) test.W(x, Y = mite.h.det, 
         MEM.autocor = "positive", 
         f = f2, 
         y = 2:10, 
         dmax = max(unlist(nbdists(x, as.matrix(mite.xy)))), 
         xy = as.matrix(mite.xy)))
# Lowest AIC, best model
mite.f2.minAIC <- 
  sapply(mite.thresh.f2, 
         function(x) min(x$best$AIC$AICc, na.rm = TRUE))
# Smallest AICc (best model among the 10)
min(mite.f2.minAIC)
# Number of the model among the 10
(nb.bestmod <- which.min(mite.f2.minAIC))
# Actual dmax of best model
(dmax.best <- mite.thresh.f2[nb.bestmod][[1]]$all[1,2])

# Extraction of the champion MEM model
mite.MEM.champ <- 
  unlist(mite.thresh.f2[which.min(mite.f2.minAIC)], 
         recursive = FALSE)
summary(mite.MEM.champ)
# Number of MEM variables in best model
(nvars.best <- which.min(mite.MEM.champ$best$AIC$AICc))
# Eigenvalues
# mite.MEM.champ$best$AIC$values # No longer in output object
# MEM variables by order of added R2
mite.MEM.champ$best$AIC$ord
# MEM variables selected in the best model
MEMid <- mite.MEM.champ$best$AIC$ord[1:nvars.best]
sort(MEMid)
MEM.all <- mite.MEM.champ$best$MEM
MEM.select <- mite.MEM.champ$best$MEM[ , sort(c(MEMid))]
colnames(MEM.select) <- sort(MEMid)
# Unadjusted R2 of best model
R2.MEMbest <- mite.MEM.champ$best$AIC$R2[nvars.best]
# Adjusted R2 of best model
RsquareAdj(R2.MEMbest, nrow(mite.h.det), length(MEMid))
# Plot the links using the function plot.links()
 
plot.links(mite.xy, thresh = dmax.best)

# RDA of the mite data constrained by the significant MEM, 
# using vegan
(mite.MEM.rda <- rda(mite.h.det~., as.data.frame(MEM.select)))
(mite.MEM.R2a <- RsquareAdj(mite.MEM.rda)$adj.r.squared)
anova(mite.MEM.rda)
(axes.MEM.test <- anova(mite.MEM.rda, by = "axis"))
# Number of significant axes
(nb.ax <- 
  length(which(axes.MEM.test[ ,ncol(axes.MEM.test)] <=  0.05)))

# Plot maps of the significant canonical axes
mite.MEM.axes <- 
  scores(mite.MEM.rda, 
         choices = 1:nb.ax, 
         display = "lc", 
         scaling = 1
)
 

if(nb.ax <=  2) 
{
     par(mfrow = c(1,2))
} else {  
     par(mfrow = c(2,2)) 
}
for(i in 1:ncol(mite.MEM.axes))
{
sr.value(mite.xy, mite.MEM.axes[ ,i])
}

# Maps of the significant MEM variables
 
if(ncol(MEM.select) <=  6) 
{
      par(mfrow = c(2,3))
}  else  { 
      par(mfrow = c(3,3)) 
}
for(i in 1:ncol(MEM.select))
{
 sr.value(mite.xy, 
          MEM.select[ ,i], 
          sub = sort(MEMid)[i], 
          csub = 2)
}

# Correlation of the retained MEM and dbMEM variables
cor(MEM.select, dbmem.red)

```


#### 7.4.3.3 其他类型的连通矩阵 


```{r}
# Other connectivity matrices

# Examples of connectivity matrices in decreasing order of
# connectivity. All these neighbourhood matrices are stored in
# objects of class nb Delaunay triangulation (as in the previous
# example)
mite.del <- tri2nb(mite.xy) 
# Gabriel graph
mite.gab <- graph2nb(gabrielneigh(as.matrix(mite.xy)), sym = TRUE)
# Relative neighbourhood
mite.rel <- graph2nb(relativeneigh(as.matrix(mite.xy)), sym = TRUE)
# Minimum spanning tree
mite.mst <- mst.nb(dist(mite.xy))

# Plots of the connectivity matrices
 
# par(mfrow = c(2,2))
plot(mite.del, mite.xy, col = "red", pch = 20, cex = 1)
title(main = "Delaunay triangulation ")
plot(mite.gab, mite.xy, col = "purple", pch = 20, cex = 1)
title(main = "Gabriel graph")
plot(mite.rel, mite.xy, col = "dark green", pch = 20, cex = 1)
title(main = "Relative neighbourhood")
plot(mite.mst, mite.xy, col = "brown", pch = 20, cex = 1)
title(main = "Minimum spanning tree")


# Link editing

# 1. Interactive:
 
plot(mite.del, mite.xy, col = "red", pch = 20, cex = 2)
title(main = "Delaunay triangulation")
mite.del2 <- edit.nb(mite.del, mite.xy)
    # To delete a link, click on its two nodes. Follow on-screen
    # instructions.
    # Wait until you have finished editing before entering the next 
    # command line. Suggestion: edit a link to site 23
    # (see below why).

# 2. Alternatively, links can also be removed by command lines, 
# after having converted the nb object into an editable matrix:
mite.del.mat <- nb2mat(mite.del, style = "B")
# Remove connection between objects 23 and 35:
mite.del.mat[23,35] <- 0
mite.del.mat[35,23] <- 0
# Back-conversion into nb object:
mite.del3 <- neig2nb(neig(mat01 = mite.del.mat))
 
plot(mite.del3, mite.xy)

# Example: list of neighbours of core #23 for the Delaunay 
#          triangulation:
mite.del[[23]]      # Before editing
mite.del2[[23]]     # After interactive editing - depends on what 
                    # you have edited above.
mite.del3[[23]]     # After command line editing


# Connectivity matrix based on a distance (radius around points)
# Using the same truncation distance dmin as in the dbMEM 
# example (1.011187).
dmin = 1.011187
mite.thresh4 <- dnearneigh(as.matrix(mite.xy), 0, dmin * 4)
# Display some values
nb2mat(mite.thresh4)[1:10,1:10]

# Using a shorter distance (1 * dmin, 2 * dmin)
mite.thresh1 <- dnearneigh(as.matrix(mite.xy), 0, dmin * 1)
mite.thresh2 <- dnearneigh(as.matrix(mite.xy), 0, dmin * 2)
# Using a longer distance
mite.thresh8 <- dnearneigh(as.matrix(mite.xy), 0, dmin * 8)

# Plot of some connectivity matrices
 
# par(mfrow = c(1,2))
plot(mite.thresh1, mite.xy, col = "red", pch = 20, cex = 0.8)
title(main = "1 * dmin")
plot(mite.thresh4, mite.xy, col = "red", pch = 20, cex = 0.8)
title(main = "4 * dmin")

# Conversion of a "nb" object into a "listw" object
# Example: mite.thresh4 created above. "B" is for "binary"
mite.thresh4.lw <- nb2listw(mite.thresh4, style = "B")
print(listw2mat(mite.thresh4.lw)[1:10,1:10], digits = 1)

# Creation of a spatial weighting matrix W = Hadamard product of 
# B and A.
# Replace "1" by Euclidean distances in the connectivity matrix
mite.thresh4.d1 <- nbdists(mite.thresh4, as.matrix(mite.xy))
# Weights as function of inverse distance
mite.inv.dist <- 
  lapply(mite.thresh4.d1, 
         function(x) 1-x/max(dist(mite.xy))
        )
# Creation of spatial weighting matrix W. Argument "B" stands for 
# "binary" but concerns the links themselves, not their weights
mite.invdist.lw <- 
  nb2listw(mite.thresh4, 
           glist = mite.inv.dist, 
           style = "B")
print(listw2mat(mite.invdist.lw)[1:10,1:10], digits = 2)

# Computation of MEM variables (from an object of class listw)
mite.invdist.MEM <- scores.listw(mite.invdist.lw)
summary(mite.invdist.MEM)
attributes(mite.invdist.MEM)$values
 
barplot(attributes(mite.invdist.MEM)$values)

# Store all MEM vectors in new object
mite.invdist.MEM.vec <- as.matrix(mite.invdist.MEM)

# Test of Moran's I of each eigenvector
mite.MEM.Moran <- 
  moran.randtest(mite.invdist.MEM.vec, mite.invdist.lw, 999)

# MEM with significant spatial correlation
which(mite.MEM.Moran$pvalue <=  0.05)
length(which(mite.MEM.Moran$pvalue <=  0.05))

# MEM with positive spatial correlation
MEM.Moran.pos <- 
  which(mite.MEM.Moran$obs > -1/(nrow(mite.invdist.MEM.vec)-1))
mite.invdist.MEM.pos <- mite.invdist.MEM.vec[ ,MEM.Moran.pos]
# MEM with positive *and significant* spatial correlation
mite.invdist.MEM.pos.sig <- 
  mite.invdist.MEM.pos[ ,which(mite.MEM.Moran$pvalue <=  0.05)]

# Plot of MEM eigenvalues vs Moran's I
 
plot(attributes(mite.invdist.MEM)$values, 
     mite.MEM.Moran$obs, 
     ylab = "Moran's I", 
     xlab = "Eigenvalues"
)
text(0, 0.55, 
     paste("Correlation = ", 
           cor(mite.MEM.Moran$obs, 
               attributes(mite.invdist.MEM)$values
               )
          )
)

```


### 7.4.4 如何选择应用正空间相关的 MEM 或者负空间相关的 MEM 

### 7.4.5 非对称的特征向量图(Asymmetric Eigenvector Maps, AEM): 方向性很重要时(When Directionality Matters)

#### 7.4.5.1 简介

#### 7.4.5.2 非对称的特征向量图(Asymmetric Eigenvector Maps, AEM)分析方法的原理与应用


```{r}

# AEM analysis ====================================================

# Coding of a river arborescence. 
# See Legendre and Legendre (2012, p. 889).
node1 <- c(1, 0, 0, 0, 0, 0, 0, 0)
n2lk6 <- c(0, 1, 0, 0, 0, 0, 0, 0)
n3lk3 <- c(1, 0, 1, 0, 0, 0, 0, 0)
n4lk2 <- c(1, 0, 0, 1, 0, 0, 0, 0)
node5 <- c(0, 1, 0, 0, 1, 0, 0, 0)
n6lk1 <- c(1, 0, 0, 1, 0, 1, 0, 0)
ln7k4 <- c(0, 1, 0, 0, 1, 0, 1, 0)
n8lk5 <- c(0, 1, 0, 0, 1, 0, 0, 1)
arbor <- rbind(node1, n2lk6, n3lk3, n4lk2, node5, 
               n6lk1, ln7k4, n8lk5)

# AEM construction
(arbor.aem <- aem(binary.mat = arbor))
arbor.aem.vec <- arbor.aem$vectors

# AEM eigenfunctions can also be obtained directly by singular 
# value decomposition (function svd()), which is what the function 
# aem() does:
arbor.c <- scale(arbor, center = TRUE, scale = FALSE)
arbor.svd <- svd(arbor.c)
# Singular values of the construction above
arbor.svd$d[1:7]
# AEM eigenfunctions of the construction above
arbor.svd$u[ ,1:7]

# Plot AEM as in book =============================================
# Coordinates of points in the figure
xx <- c(2.75, 2, 3.5, 1, 2, 3.5, 2, 3, 4)
yy <- c(1,2,2,3,3,3,4,4,4)
data.xy <- data.frame(xx,yy)
rownames(data.xy) <- 
  c("N0", "N1", "N2", "N3", "N4", "N5", "N6", "N7", "N8")
arbor.points <- 
  rbind(rep(0,7), arbor.svd$u[ ,1:7]) # With node 0 added
# Change signs of AEM 5
arbor.points[ ,5] <- -arbor.points[ ,5]
 
# par(mfrow = c(3,3))
for(i in 1:7) {
sr.value(data.xy, 
         arbor.points[ ,i], 
         method = "greylevel", 
         csize = 1, 
         clegend = 0, 
         grid = FALSE, 
         include.origin = FALSE, 
         addaxes = FALSE, 
         sub = paste("AEM",i), 
         csub = 2)
}
# END Plot AEM as in book =========================================

# Coding of sampling design: 10 cross-river transects, 4 traps 
# per transect. Edges weighted proportional to inverse squared
# distance.
# X-Y coordinates
xy <- cbind(1:40, expand.grid(1:4, 1:10))
# Object of class nb (spdep) containing links of chess type "queen"
nb <- cell2nb(4, 10, "queen")
# Site-by-edges matrix (produces a fictitious object "0")
# with automatic plot of the site-by-edges matrix

edge.mat <- aem.build.binary(nb, xy)
# Matrix of Euclidean distances
D1.mat <- as.matrix(dist(xy))
# Extract the edges, remove the ones directly linked to site 0
edges.b <- edge.mat$edges[-1:-4,]
# Construct a vector giving the length of each edge
length.edge <- vector(length = nrow(edges.b))
for(i in 1:nrow(edges.b))
{
      length.edge[i] <- D1.mat[edges.b[i,1], edges.b[i,2]]
}
# Weighting of edges based on inverse squared distance
weight.vec <- 1-(length.edge/max(length.edge))^2
# Construction of AEM eigenfunctions from edge.mat, of class
# build.binary
example.AEM <- 
  aem(aem.build.binary = edge.mat, 
      weight = weight.vec, 
      rm.link0 = TRUE)
example.AEM$values
ex.AEM.vec <- example.AEM$vectors

# Construction of 5 fictitious species
# Two randomly distributed species
sp12 <- matrix(trunc(rnorm(80,5,2),0),40)
# One species restricted to the upper half of the stream
sp3 <- c(trunc(rnorm(20,8,2.5),0), rep(0,20))
# One species restricted to the left-hand half of the transect
sp4 <- t(matrix(c(trunc(rnorm(20,8,3),0), rep(0,20)),10))
sp4 <- as.vector(sp4)
# One species restricted to the 4 upper left-hand sites
sp5 <- c(4,7,0,0,3,8, rep(0,34))
# Build the species matrix
sp <- cbind(sp12, sp3, sp4, sp5)
colnames(sp) <- c("sp1", "sp2", "sp3", "sp4", "sp5")

# Global AEM analysis with 20 first AEM variables (for 
# computation of R2a)
AEM.20 <- rda(sp ~ ., as.data.frame(ex.AEM.vec[ ,1:20]))
(R2a.AEM <- RsquareAdj(AEM.20)$adj.r.squared)

# Forward selection of the AEM variables
AEM.fwd <- forward.sel(sp, ex.AEM.vec, adjR2thresh = R2a.AEM)
(AEM.sign <- sort(AEM.fwd[ ,2]))
# Write significant AEM in a new object
AEM.sign.vec <- ex.AEM.vec[ ,c(AEM.sign)]
# RDA with signif. AEM
(sp.AEMsign.rda <- rda(sp ~ ., data = as.data.frame(AEM.sign.vec)))
anova(sp.AEMsign.rda)
(AEM.rda.axes.test <- anova(sp.AEMsign.rda, by = "axis"))
# Number of significant axes
(nb.ax.AEM <- length(which(AEM.rda.axes.test[ ,4] <=  0.05)))
# Adjusted R-square
RsquareAdj(sp.AEMsign.rda)

# Plot of the significant canonical axes
AEM.rda.axes <- 
  scores(sp.AEMsign.rda, 
         choices = c(1,2), 
         display = "lc", 
         scaling = 1)
 
par(mfrow = c(1,nb.ax.AEM))
for(i in 1:nb.ax.AEM) sr.value(xy[ ,c(2,3)], AEM.rda.axes[ ,i])



```


---

## 7.5 另一种空间结构的观察方法: 多尺度排序(Multiscale Ordination, MSO)

### 7.5.1 多尺度排序(Multiscale Ordination, MSO)方法的基本原理

### 7.5.2 多尺度排序(Multiscale Ordination, MSO)方法在甲螨数据集探索性分析中的应用


```{r}
## MSO of the undetrended mite data vs environment RDA
mite.undet.env.rda <- rda(mite.h~., mite.env2)
(mite.env.rda.mso <- 
  mso(mite.undet.env.rda, 
      mite.xy, 
      grain = dmin, 
      perm = 999))
 
msoplot(mite.env.rda.mso, alpha = 0.05/7)

# MSO of the undetrended mite data vs environment RDA, controlling 
# for MEM
mite.undet.env.MEM <- 
  rda(mite.h, 
      mite.env2, 
      as.data.frame(MEM.select))
(mite.env.MEM.mso <- 
  mso(mite.undet.env.MEM, 
      mite.xy, 
      grain = dmin, 
      perm = 999))
 
msoplot(mite.env.MEM.mso, alpha = 0.05/7) 
# msoplot(mite.env.MEM.mso, 
#        alpha = 0.05/7, 
#        ylim = c(0, 0.0045)  # Expanded height to clear legend
# )
```


### 7.5.3 多尺度排序(Multiscale Ordination, MSO)方法在甲螨数据集和环境数据集中的应用

```{r}
## MSO on detrended mite and environmental data
# Detrend mite data on Y coordinate
mite.h.det2 <- resid(lm(as.matrix(mite.h) ~ mite.xy[ ,2]))
# Detrend environmental data on Y coordinate
env2.det <- as.data.frame(
                    resid(lm(as.matrix(mite.env2) ~ mite.xy[ ,2])))
# RDA and MSO
mitedet.envdet.rda <- rda(mite.h.det2 ~., env2.det)
(miteenvdet.rda.mso <- 
  mso(mitedet.envdet.rda, 
      mite.xy, 
      grain = dmin, 
      perm = 999))
 
msoplot(miteenvdet.rda.mso, alpha = 0.05/7)
# msoplot(miteenvdet.rda.mso, alpha = 0.05/7, ylim = c(0, 0.006))

## MSO of the detrended mite data vs detrended environment RDA, 
## controlling for MEM
mite.det.env.MEM <- 
  rda(mite.h.det2, 
      env2.det, 
      as.data.frame(MEM.select))
(mite.env.MEM.mso <- 
  mso(mite.det.env.MEM, 
      mite.xy, 
      grain = dmin, 
      perm = 999))
 
msoplot(mite.env.MEM.mso, alpha = 0.05/7)
# msoplot(mite.env.MEM.mso, alpha = 0.05/7, ylim = c(0, 0.005))
```


---

## 7.6 多元变量方差分析中的无重复时空交互检验(Space-Time Interaction Test in Multivariate ANOVA, Without Replicates)

### 7.6.1 简介

### 7.6.2 使用 sti() 函数进行时空交互检验



```{r}
# Space-time interaction analysis of the Trichoptera data =========

# Load the Trichoptera data
data(trichoptera)
names(trichoptera) # Species codes

# Hellinger transformation
tricho.hel <- decostand(trichoptera[ ,-c(1,2)], "hel")

# sti analysis
stimodels(tricho.hel, S = 22, Ti = 10, model = "5")

# Quick and easy sti analysis using function quicksti()
quicksti(tricho.hel, S = 22, Ti = 10)
```


---

## 7.7 结论



---

</font>
