<font face="微软雅黑">

---

# **第八章：群落多样性分析**

---

本章内容简介：

- 8.1 目标
- 8.2 关于多样性的介绍
- 8.3 与空间相关的多样性: Alpha, Beta and Gamma 多样性
- 8.4 β-多样性(Beta Diversity)
- 8.5 群落的功能性多样性 & 功能性组成 & 系统进化多样性 
- 8.6 结论

---

## 8.1 目标



---

## 8.2 关于多样性的介绍



### 8.2.1 简介

### 8.2.2 简单数字统计的物种多样性

#### 8.2.2.1 物种的丰富度与稀疏性(Species Richness & Rarefaction)

#### 8.2.2.2 物种丰度多样性的组成: 丰富度与均匀度(Species Abundance Diversity Components: Richness and Evenness)

### 8.2.3 分类学多样性指标的实际应用(Taxonomic Diversity Indices)

#### 8.2.3.1 河流数据集中鱼类群落的物种多样性指标(Species Diversity Indices)


```{r}
# Load packages, functions and data ===============================
library(ade4)
library(adegraphics)
library(adespatial)
library(vegan)
library(vegetarian)
library(ggplot2)
library(FD)
library(taxize)

# Source additional functions that will be used later in this
# Chapter. Our scripts assume that files to be read are in
# the working directory.
source("Functions/panelutils.R")
source("Functions/Rao.R")

# Load the Doubs data. The file Doubs.Rdata is assumed to be in 
# the working directory
load("Data/Doubs.RData") 
load("Data/mite.RData")
# Remove empty site 8
spe <- spe[-8, ]
env <- env[-8, ]
spa <- spa[-8, ]

# Load the oribatid mite data. The file mite.Rdata is assumed 
# to be in the working directory.




```



```{r}
# Alpha species diversity =========================================

# Get help on the diversity() function
# ?diversity

# Compute alpha diversity indices of the fish communities
N0 <- rowSums(spe > 0)          # Species richness
N0 <- specnumber(spe)           # Species richness (alternate)
H <- diversity(spe)             # Shannon entropy (base e)
Hb2 <- diversity(spe, base = 2) # Shannon entropy (base 2)
N1 <- exp(H)                    # Shannon diversity (base e)
                                # (number of abundant species)
N1b2 <- 2^Hb2                   # Shannon diversity (base 2)
N2 <- diversity(spe, "inv")     # Simpson diversity 
                                # (number of dominant species)
J <- H / log(N0)                # Pielou evenness
E10 <- N1 / N0                  # Shannon evenness (Hill's ratio)
E20 <- N2 / N0                  # Simpson evenness (Hill's ratio)
(div <- data.frame(N0, H, Hb2, N1, N1b2, N2, E10, E20, J))

# Correlations among diversity indices
cor(div)

 
pairs(div[-1, ], 
      lower.panel = panel.smooth, 
      upper.panel = panel.cor,
      diag.panel = panel.hist, main = "Pearson Correlation Matrix"
)
```


#### 8.2.3.2 甲螨群落的稀疏性分析(Rarefaction Analysis)


```{r}
# Rarefaction analysis
# Number of species in the 70 moss or soil cores
(mite.nbsp <- specnumber(mite))
# Cores with minimum and maximum observed species richness
mite.nbsp[mite.nbsp == min(mite.nbsp)]
mite.nbsp[mite.nbsp == max(mite.nbsp)]
range(mite.nbsp)
# Total abundance in the 70 cores
(mite.abund <- rowSums(mite))
range(mite.abund)
# Abundance in the cores with smallest number of species
mite.abund[mite.nbsp == min(mite.nbsp)]
# Abundance in the core with largest number of species
mite.abund[mite.nbsp == max(mite.nbsp)]
# Number of species in the core with smallest abundance
mite.nbsp[mite.abund == min(mite.abund)]
# Number of species in the core with largest abundance
mite.nbsp[mite.abund == max(mite.abund)]

# Rarefaction to 80 individuals
mite.rare80 <- rarefy(mite, sample = 80)
# Compare ranking of observed and rarefied cores
sort(mite.nbsp)
sort(round(mite.rare80))
# Sites with minimum and maximum estimated species richness
mite.rare80[mite.rare80 == min(mite.rare80)]
mite.rare80[mite.rare80 == max(mite.rare80)]
# Observed core with smallest predicted species richness
mite[which(mite.rare80 == min(mite.rare80)),]
# Observed core with largest predicted species richness
mite[which(mite.rare80 == max(mite.rare80)),]

# Rarefaction curve
 
rarecurve(
  mite[-67,], 
  step = 1, 
  sample = 80, 
  xlab = "Number of individuals (Sample Size)", 
  ylab = "Species",
  label = TRUE, 
  col = "blue"
)

# Same plot but including core #67:
# rarecurve(
#   mite, 
#   step = 1, 
#   sample = 80, 
#   xlab = "Number of individuals (Sample Size)", 
#   ylab = "Species",
#   label = TRUE, 
#   col = "blue"
# )

```



---

## 8.3 与空间相关的多样性: Alpha, Beta and Gamma 多样性



---

## 8.4 β-多样性(Beta Diversity)

### 8.4.1 简单数字统计的β-多样性


```{r}

# Beta diversity, LCBD and SCBD of the fish data ==================

# Gamma richness and expected species pool
# ?specpool
(gobs <- ncol(spe))
(gthe <- specpool(spe))


# Multiplicative partitioning of Hill numbers (Jost 2006, 2007)
# ?d
# Mean alpha species richness
d(spe, lev = "alpha", q = 0)
# Mean alpha Shannon diversity
d(spe, lev = "alpha", q = 1)
# Mean alpha Simpson diversity
d(spe, lev = "alpha", q = 2, boot = TRUE)

# Multiplicative beta species richness
d(spe, lev = "beta", q = 0)
# Multiplicative beta Shannon diversity
d(spe, lev = "beta", q = 1)
# Multiplicative beta Simpson diversity
d(spe, lev = "beta", q = 2, boot = TRUE)

# Gamma species richness
d(spe, lev = "gamma", q = 0)
# Gamma Shannon diversity
d(spe, lev = "gamma", q = 1)
# Gamma Simpson diversity
d(spe, lev = "gamma", q = 2, boot = TRUE)

# Plot multiplicative beta diversity vs order
mbeta <- data.frame(order = 0:20, beta = NA, se = NA)
for (i in 1:nrow(mbeta)) {
  out <- d(spe, lev = "beta", q = mbeta$order[i], boot = TRUE)
  mbeta$beta[i] <- out$D.Value
  mbeta$se[i] <- out$StdErr
}
mbeta
 
ggplot(mbeta, aes(order, beta)) +
  geom_point() +
  geom_line() +
  geom_errorbar(
     aes(order, beta, ymin = beta - se, ymax = beta + se),
     width = 0.2) + labs(y = "Multiplicative beta diversity", 
                         x = "Order of the diversity measure")


# MacArthur's homogeneity measure (MacArthur 1965)
hom <- data.frame(order = 0:20, homogeneity = NA, se = NA)
for (i in 1:nrow(hom)) {
  out <- M.homog(spe, q = hom$order[i], std = TRUE, boot = TRUE)
  hom$homogeneity[i] <- out$M
  hom$se[i] <- out$StdErr
}
hom
 
ggplot(hom, aes(order, homogeneity)) +
  geom_point() +
  geom_line() +
  geom_errorbar(
     aes(order, 
           homogeneity, 
           ymin = homogeneity - se, 
           ymax = homogeneity + se),
         width = 0.2) + labs(y = "MacArthur's homogeneity measure",
                             x = "Order of the diversity measure")

```


### 8.4.2 将 β-多样性作为群落组成表中的变量: SCBD & LCBD 指标

#### 8.4.2.1 简介

#### 8.4.2.2 计算 β-多样性

#### 8.4.2.3 关于β-多样性的相异性度量的性质(Properties of Dissimilarity Measures with Respect to Beta Diversity)

#### 8.4.2.4 β-多样性的局部贡献(Local Contributions to Beta Diversity, LCBD)

#### 8.4.2.5 β-多样性的物种贡献(Species Contributions to Beta Diversity, SCBD)

#### 8.4.2.6 使用 adespatial 分析包中的 beta.div() 函数计算 LCBD & SCBD 


```{r}
# Computation using beta.div {adespatial} on 
# Hellinger-transformed species data
spe.beta <- beta.div(spe, method = "hellinger", nperm = 9999)
summary(spe.beta)
spe.beta$beta  # SSTotal and BDTotal

# Which species have a SCBD larger than the mean SCBD?
spe.beta$SCBD[spe.beta$SCBD >= mean(spe.beta$SCBD)]

```


```{r}
### NOT IN THE BOOK ###

# Plot of the species with large SCBD
 
par(mfrow = c(2,3))
plot(spa, 
     asp = 1, 
     cex.axis = 0.8, 
     col = "brown", 
     cex = spe$Satr, 
     main = "Brown trout", 
     xlab = "x coordinate (km)", 
     ylab = "y coordinate (km)"
)
lines(spa, col = "light blue")
plot(spa, 
     asp = 1, 
     cex.axis = 0.8, 
     col = "brown", 
     cex = spe$Phph, 
     main = "Eurasian minnow", 
     xlab = "x coordinate (km)", 
     ylab = "y coordinate (km)"
)
lines(spa, col = "light blue")
plot(spa, 
     asp = 1, 
     cex.axis = 0.8, 
     col = "brown", 
     cex = spe$Babl, 
     main = "Stone loach", 
     xlab = "x coordinate (km)", 
     ylab = "y coordinate (km)"
)
lines(spa, col = "light blue")
plot(spa, 
     asp = 1, 
     cex.axis = 0.8, 
     col = "brown", 
     cex = spe$Ruru, 
     main = "Roach", 
     xlab = "x coordinate (km)", 
     ylab = "y coordinate (km)"
)
lines(spa, col = "light blue")
plot(spa, 
     asp = 1, 
     cex.axis = 0.8, 
     col = "brown", 
     cex = spe$Alal, 
     main = "Bleak", 
     xlab = "x coordinate (km)", 
     ylab = "y coordinate (km)"
)
lines(spa, col = "light blue")

### END NOT IN THE BOOK ###
```



```{r}
# LCBD values
spe.beta$LCBD
# p-values
spe.beta$p.LCBD
# Holm correction
p.adjust(spe.beta$p.LCBD, "holm")
# Sites with significant Holm-corrected LCBD value
row.names(spe[which(p.adjust(spe.beta$p.LCBD, "holm") <= 0.05),])

# Plot the LCBD values on the river map
 
plot(spa, 
     asp = 1, 
     cex.axis = 0.8, 
     pch = 21, 
     col = "white", 
     bg = "brown", 
     cex = spe.beta$LCBD * 70, 
     main = "LCBD values", 
     xlab = "x coordinate (km)", 
     ylab = "y coordinate (km)"
)
lines(spa, col = "light blue")
text(85, 11, "***", cex = 1.2, col = "red")
text(80, 92, "***", cex = 1.2, col = "red")

```


### 8.4.3 将β-多样性分解成替换 & 丰富度差异 & 嵌套组成(Replacement, Richness Difference and Nestedness Components)

#### 8.4.3.1 Podani 指数(Podani Family of Indices)

#### 8.4.3.2 Baselga 指数(Baselga Family of Indices)

#### 8.4.3.3 嵌套介绍(Flavours of Nestedness)

#### 8.4.3.4 使用 adespatial 分析包中的 beta.div.comp() 函数计算替换 & 丰富度差异 & 嵌套组成(Replacement & Richness Difference & Nestedness)


```{r}

# Replacement, richness and nestedness indices of the fish data ===

# Jaccard-based Podani indices (presence-absence data)
fish.pod.j <- beta.div.comp(spe, coef = "J", quant = FALSE)
# What is in the output object?
summary(fish.pod.j)
# Display summary statistics:
fish.pod.j$part
# Extraction of the richness difference matrix
fish.rich <- as.matrix(fish.pod.j$rich)
# Plot of the richness difference with respect to site 30
fish.rich.30 <- fish.rich[29, ][-29]
site.names <- seq(1, 29)[-8]
 
plot(site.names, 
  fish.rich.30, 
  type = "n", 
  xaxp = c(1, 29, 28),
  main = "Doubs fish data: richness difference with respect to site 30", 
  xlab = "Site number", 
  ylab = "Richness difference"
)
lines(site.names, fish.rich.30, pch = 24, col = "red")
points(
  site.names, 
  fish.rich.30, 
  pch = 24, 
  cex = 1.5, 
  col = "white", 
  bg = "red"
)
text(3, 0.85, "Upstream", cex = 1, col = "red")
text(27, 0.85, "Downstream", cex = 1, col = "red")

# Extraction of the replacement matrix
fish.repl <- as.matrix(fish.pod.j$repl)
# Extraction of the Jaccard dissimilarity Dj matrix
fish.jac <- as.matrix(fish.pod.j$D)

# Plot of the Jaccard, replacement and richness difference indices 
# between nearest neighbours
# First, extract the subdiagonals of the square dissimilarity 
# matrices
fish.repl.neigh <- diag(fish.repl[-1, ]) # Replacement
fish.rich.neigh <- diag(fish.rich[-1, ]) # Richness difference
fish.jac.neigh <- diag(fish.jac[-1, ]) # Jaccard Dj index

 
absc <- c(2:7, 9:30) # Abscissa
label.pairs <- c("1-2", "2-3", "3-4", "4-5", "5-6", "6-7", " ", 
  "7-9", "9-10", "10-11", "11-12", "12-13", "13-14", "14-15", 
  "15-16", "16-17", "17-18", "18-19", "19-20", "20-21", "21-22", 
  "22-23", "23-24", "24-25", "25-26", "26-27", "27-28", "28-29", 
  "29-30")
plot(
  absc, 
  fish.jac.neigh, 
  type = "n", 
  xaxt = "n",
  main = "Replacement - Richness difference - Jaccard - nearest neighbours",  
  xlab = "Site pairs", 
  ylab = "Podani's indices"
)
axis(side = 1, 2:30, labels = label.pairs, las = 2, cex.axis = 0.9)
lines(absc, fish.jac.neigh, col = "black")
points(
  absc, 
  fish.jac.neigh, 
  pch = 21, 
  cex = 2, 
  col = "black", 
  bg = "black"
)
lines(absc, fish.repl.neigh, col = "blue")
points(
  absc, 
  fish.repl.neigh, 
  pch = 22, 
  cex = 2, 
  col = "white", 
  bg = "blue"
)
lines(absc, fish.rich.neigh, col = "red")
points(
  absc, 
  fish.rich.neigh, 
  pch = 24, 
  cex = 2, 
  col = "white", 
  bg = "red"
)
legend(
  "top", 
  c("Jaccard D", "Replacement", "Richness difference"), 
  pch = c(16, 15, 17), 
  col = c("black", "blue", "red")
)

# Triangular plots
# Jaccard
fish.pod.J <- beta.div.comp(spe, coef = "J", quant = FALSE)
# Sorensen
fish.pod.S <- beta.div.comp(spe, coef = "S", quant = FALSE)
# Ruzicka
fish.pod.qJ <- beta.div.comp(spe, coef = "J", quant = TRUE)
# Percentage difference
fish.pod.qS <- beta.div.comp(spe, coef = "S", quant = TRUE)
# Data frames for the triangular plots
fish.pod.J.3 <- cbind((1 - fish.pod.J$D),
                      fish.pod.J$repl,
                      fish.pod.J$rich)
colnames(fish.pod.J.3) <- c("Similarity", "Repl", "RichDiff")
fish.pod.S.3 <- cbind((1 - fish.pod.S$D),
                      fish.pod.S$repl,
                      fish.pod.S$rich)
colnames(fish.pod.S.3) <- c("Similarity", "Repl", "RichDiff")
fish.pod.qJ.3 <- cbind((1 - fish.pod.qJ$D),
                       fish.pod.qJ$repl,
                       fish.pod.qJ$rich)
colnames(fish.pod.qJ.3) <- c("Similarity", "Repl", "AbDiff")
fish.pod.qS.3 <- cbind((1 - fish.pod.qS$D),
                       fish.pod.qS$repl,
                       fish.pod.qS$rich)
colnames(fish.pod.qS.3) <- c("Similarity", "Repl", "AbDiff")

 
par(mfrow = c(2, 2))
triangle.plot(
  as.data.frame(fish.pod.J.3[, c(3, 1, 2)]),
  show = FALSE,
  labeltriangle = FALSE,
  addmean = TRUE
)
text(-0.45, 0.5, "RichDiff", cex = 1.5)
text(0.4, 0.5, "Repl", cex = 1.5)
text(0, -0.6, "Jaccard similarity", cex = 1.5)
triangle.plot(
  as.data.frame(fish.pod.S.3[, c(3, 1, 2)]),
  show = FALSE,
  labeltriangle = FALSE,
  addmean = TRUE
)
text(-0.45, 0.5, "RichDiff", cex = 1.5)
text(0.4, 0.5, "Repl", cex = 1.5)
text(0, -0.6, "Sørensen similarity", cex = 1.5)

triangle.plot(
  as.data.frame(fish.pod.qJ.3[, c(3, 1, 2)]),
  show = FALSE,
  labeltriangle = FALSE,
  addmean = TRUE
)
text(-0.45, 0.5, "AbDiff", cex = 1.5)
text(0.4, 0.5, "Repl", cex = 1.5)
text(0, -0.6, "S = 1 – Ružička D", cex = 1.5)
triangle.plot(
  as.data.frame(fish.pod.qS.3[, c(3, 1, 2)]),
  show = FALSE,
  labeltriangle = FALSE,
  addmean = TRUE
)
text(-0.45, 0.5, "AbDiff", cex = 1.5)
text(0.4, 0.5, "Repl", cex = 1.5)
text(0, -0.6, "S = 1 – Percentage difference", cex = 1.5)
# Display values of the mean points in the triangular plots
colMeans(fish.pod.J.3[, c(3, 1, 2)])
colMeans(fish.pod.S.3[, c(3, 1, 2)])
colMeans(fish.pod.qJ.3[, c(3, 1, 2)])
colMeans(fish.pod.qS.3[, c(3, 1, 2)])

```


#### 8.4.3.5 解释性替代与丰富度差异(Explaining Replacement and Richness Difference)

```{r}
# Replacement
repl.dbrda <- dbrda(fish.repl ~ ., data = env, add = "cailliez")
anova(repl.dbrda)
RsquareAdj(repl.dbrda)

# Richness difference
rich.dbrda <- dbrda(fish.rich ~ ., data = env, add = "cailliez")
anova(rich.dbrda)
RsquareAdj(rich.dbrda)

 
plot(
  rich.dbrda,
  scaling = 1,
  display = c("lc", "cn"),
  main = "Richness difference explained by environmental variables"
)
```


---

## 8.5 群落的功能性多样性 & 功能性组成 & 系统进化多样性

### 8.5.1 α-功能多样性(Alpha Functional Diversity)



```{r}
summary(fishtraits)
rownames(fishtraits)
names(spe)
names(fishtraits)
tra <- fishtraits[ , 6:15]
tra 

# Distance-based functional diversity indices
# ?dbFD
 

# In the code below to create object res, run the function and
# then answer the questions as suggested below,
# i.e., g then 10.

res <-
  dbFD(
    tra,
    as.matrix(spe),
    asym.bin = 5:10,
    stand.FRic = TRUE,
    clust.type = "ward.D",
    CWM.type = "all",
    calc.FGR = TRUE
  )
# g  # cut the dendrogram using the number of groups as criterion
# 10 #<<<<< choose the number of functional groups
res

 
par(mfrow = c(1, 2))
plot(spa, 
     asp = 1, 
     pch = 21, 
     cex.axis = 0.8, 
     col = "white", 
     bg = "brown", 
     cex = res$FRic * 5, 
     main = "Functional richness", 
     xlab = "x coordinate (km)", 
     ylab = "y coordinate (km)"
)
lines(spa, col = "light blue")
plot(spa, 
     asp = 1, 
     pch = 21, 
     cex.axis = 0.8, 
     col = "white", 
     bg = "brown", 
     cex = res$FEve * 6, 
     main = "Functional evenness", 
     xlab = "x coordinate (km)", 
     ylab = "y coordinate (km)"
)
lines(spa, col = "light blue")

plot(spa, 
     asp = 1, 
     pch = 21, 
     cex.axis = 0.8, 
     col = "white", 
     bg = "brown", 
     cex = res$FDiv * 5, 
     main = "Functional divergence", 
     xlab = "x coordinate (km)", 
     ylab = "y coordinate (km)"
)
lines(spa, col = "light blue")
plot(spa, 
     asp = 1, 
     pch = 21, 
     cex.axis = 0.8, 
     col = "white", 
     bg = "brown", 
     cex = res$FDis * 1.5, 
     main = "Functional dispersion", 
     xlab = "x coordinate (km)", 
     ylab = "y coordinate (km)"
)
lines(spa, col = "light blue")

plot(spa, 
     asp = 1, 
     pch = 21, 
     cex.axis = 0.8, 
     col = "white", 
     bg = "brown", 
     cex = res$RaoQ / 2, 
     main = "Rao quadratic entropy", 
     xlab = "x coordinate (km)", 
     ylab = "y coordinate (km)"
)
lines(spa, col = "light blue")
plot(spa, 
     asp = 1, 
     pch = 21, 
     cex.axis = 0.8, 
     col = "white", 
     bg = "brown", 
     cex = res$FGR / 2, 
     main = "Functional group richness", 
     xlab = "x coordinate (km)", 
     ylab = "y coordinate (km)"
)
lines(spa, col = "light blue")

# Add these indices to the div data frame
div$FRic <- res$FRic
div$FEve <- res$FEve
div$FDiv <- res$FDiv
div$FDis <- res$FDis
div$RaoQ <- res$RaoQ
div$FGR <- res$FGR
div

# Community-weighted mean trait values (CWMs)
functcomp(tra, as.matrix(spe), CWM.type = "all")

```


### 8.5.2 β 分类学 & 系统发育学 & 功能性多样性(Beta Taxonomic, Phylogenetic and Functional Diversities)

```{r}

# Distance matrix based on a simplified phylogenetic classification
# Retreive hierarchical classification from species list
splist <- as.character(fishtraits$LatinName)
# Database ncbi - less reliable than GBIF
spcla <- classification(splist, db = "ncbi")
# spcla <- classification(splist, db = "gbif")

# ------------------------------------------------------------
# WARNING: depending on the taxonomic status of the species,
# the function classification() may stop and ask the user to choose 
# between two (or more) names. However, if this script is copied 
# and pasted as a whole, then NAs will be produced and some species
# will be deleted. This will induce errors in the next steps of
# the analysis. To avoid that, run the "spcal <- classification..."
# only and make your choice. The function will then proceed.
# ------------------------------------------------------------


# Compute the distance matrix and the phylogenetic tree
tr <- class2tree(spcla)
tr$classification
tr$distmat
# Convert the tree to a cophenetic matrix
# constrained between 0 and 1
phylo.d <- cophenetic(tr$phylo) / 100
# Replace full species names by name codes
rownames(phylo.d) <- names(spe)
colnames(phylo.d) <- names(spe)

# Functional dissimilarity matrix (Gower dissimilarity)
trait.d <- gowdis(tra, asym.bin = 5:10)

# Plot the tree and the dendrogram 
trait.gw <- hclust(trait.d, "ward.D2")
 
# par(mfrow = c(1, 2))
plot(tr)
text(0.1, 27.5, "a", cex = 1.8)
plot(trait.gw, hang = -1, main = "")
text(26, 1.045, "b", cex = 1.8)

# Additive partitioning of TD, FD and PD ######
# (de Bello et al. 2010)
spe.rao <- Rao(
  sample = t(spe),
  dfunc = trait.d, # optional functional dissimilarity matrix
  dphyl = phylo.d, # optional phylogenetic distance matrix
  weight = FALSE,
  Jost = TRUE,
  structure = NULL
)
names(spe.rao)

# Species diversity (Simpson)
names(spe.rao$TD)
# Mean alpha Simpson diversity
spe.rao$TD$Mean_Alpha
# Gamma Simpson diversity
spe.rao$TD$Gamma
# Additive beta Simpson diversity
spe.rao$TD$Beta_add
spe.rao$TD$Gamma - spe.rao$TD$Mean_Alpha
# Beta diversity expressed as percentage of gamma
spe.rao$TD$Beta_prop
spe.rao$TD$Beta_add / spe.rao$TD$Gamma
# Multiplicative beta Simpson diversity
spe.rao$TD$Gamma / spe.rao$TD$Mean_Alpha

# Phylogenetic diversity (Rao)
names(spe.rao$PD)
# Mean alpha Rao phylogenetic diversity
spe.rao$PD$Mean_Alpha
# Gamma Rao phylogenetic diversity
spe.rao$PD$Gamma
# Additive beta Rao phylogenetic diversity
spe.rao$PD$Beta_add
spe.rao$PD$Gamma - spe.rao$PD$Mean_Alpha
# Beta phylogenetic diversity expressed as percentage of gamma
spe.rao$PD$Beta_prop
spe.rao$PD$Beta_add / spe.rao$PD$Gamma
# Multiplicative beta Rao phylogenetic diversity
spe.rao$PD$Gamma / spe.rao$PD$Mean_Alpha

# Functional diversity (Rao)
names(spe.rao$FD)
# Mean alpha Rao functional diversity
spe.rao$FD$Mean_Alpha
# Gamma Rao functional diversity
spe.rao$FD$Gamma
# Additive beta Rao functional diversity
spe.rao$FD$Beta_add
spe.rao$FD$Gamma - spe.rao$FD$Mean_Alpha
# Beta functional diversity expressed as percentage of gamma
spe.rao$FD$Beta_prop
spe.rao$FD$Beta_add / spe.rao$FD$Gamma
# Multiplicative beta Rao functional diversity
spe.rao$FD$Gamma / spe.rao$FD$Mean_Alpha

# Variation of alpha TD, PD and FD along the Doubs river
spe.rao$TD$Alpha
spe.rao$PD$Alpha
spe.rao$FD$Alpha

 
par(mfrow = c(2, 2))
plot(spa, 
     asp = 1, 
     cex.axis = 0.8, 
     pch = 21, 
     col = "white", 
     bg = "brown", 
     cex = spe.rao$TD$Alpha / 4, 
     main = "Taxonomic diversity", 
     xlab = "x coordinate (km)", 
     ylab = "y coordinate (km)"
)
lines(spa, col = "light blue")
plot(spa, 
     asp = 1, 
     cex.axis = 0.8, 
     pch = 21, 
     col = "white", 
     bg = "brown", 
     cex = spe.rao$PD$Alpha * 1.5, 
     main = "Phylogenetic diversity", 
     xlab = "x coordinate (km)", 
     ylab = "y coordinate (km)"
)
lines(spa, col = "light blue")

plot(spa, 
     asp = 1, 
     cex.axis = 0.8, 
     pch = 21, 
     col = "white", 
     bg = "brown", 
     cex = spe.rao$FD$Alpha * 2.5, 
     main = "Functional diversity", 
     xlab = "x coordinate (km)", 
     ylab = "y coordinate (km)"
)
lines(spa, col = "light blue")

# Add Rao-based diversity indices to the div data frame
div$alphaTD <- spe.rao$TD$Alpha
div$alphaPD <- spe.rao$PD$Alpha
div$alphaFD <- spe.rao$FD$Alpha

# Save the div data frame as a CSV file
write.csv(div, file = "diversity.csv", quote = FALSE)

# Save the species classification as a CSV file
write.csv(tr$classification, file = "classification.csv", 
   quote = FALSE)
```


---

## 8.6 结论



---

</font>
